### 2021-11-10 NSF2021/Description/specificaim2.tex

\begin{wrapfigure}[]{l}{1.8in}
  \centering
  \input{figures/doublewell.tikz}
  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
\end{wrapfigure}
In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
\begin{wrapfigure}[]{r}{1.8in}
  \centering
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b =
  0.4$.}
\end{wrapfigure}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve).

Solutions of the screened Poisson equation can be written
in terms of volume and layer potentials. We first define the level set
$\Gamma = \{\xx \:|\: \uu(\xx) = (U_0 + U_1)/2 \}$ which separates the
domain $\Omega$ into two regions. Then, assuming that the particles are
all contained in the hydrophobic region, a solution of the screened
Laplace equation (i.e.~equation~\eqref{eq:screenedPoisson} when $F[u^N]
= 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}

>>>>

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha (dF)^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''$ is the Fr\'{e}chet derivative of $F$,
and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
\begin{wrapfigure}[]{l}{1.8in}
  \centering
  \input{figures/doublewell.tikz}
  \caption{\label{fig:well} (a) The double well potential when $U_1 =
  0.5$, $U_2 = 2$, $a = 1$, and $b = 0.4$. (b) The corresponding
  solution of~\eqref{eq:F}.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve). As expected, the exact Newton's requires less
iterations, but applying the inexact Newton's method is much cheaper
since the $\tilde{f}''$ is piecewise constant.
\begin{wrapfigure}[]{r}{1.8in}
  \centering
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} The max norm of the residual of
  inexact (blue) and exact (red) Newton's method.}
\end{wrapfigure}

Having demonstrated that the inexact Newton method is a viable solver
for the one-dimensional Cahn-Allen equation, we propose to extend the
results to higher spatial dimensions. Again, the biggest challenge is
solving the piecewise constant PDE~\eqref{eq:screenedPoisson}. To
exploit the benefits of an integral equation methods, we will write the
solution of~\eqref{eq:screenedPoisson} in terms of layer and volume
potentials. We start by defining the level set $\Gamma = \{\xx \:|\:
\uu(\xx) = (U_0 + U_1)/2 \}$ which separates the domain $\Omega$ into
the hydrophobic order region and the bulk order region. In each of these
regions $f''(u)$ is constant. Then
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}
